Continous time Markov chain models; \\ extension, simulation and visualisation

### 2014-10-21 chapters/chapter01.tex
working on plots

%PL: Not sure if you want to do a color page print for this figure; it could simply be black/white

%---SUBSTITUTION MODEL---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{substitution} 
\caption{
{ \footnotesize 
{\bf  Graphical representation of  a nucleotide substitution model.} A Markov substitution model with 6 parameters describing the symmetric transition rates.
} % END: footnotesize
}
\label{fig:substitution}
\end{figure}

@@

%TODO: JK69 as example
\subsection{A simple example: the Jukes and Cantor substitution model\label{sub:jc69}}

@@

\noindent
Second, the rate and time in Equation~(\ref{eq:jc69Finite}) occur as a product, which means that rate and time are confounded.
If we were to simulate two sequences, using the algorithm listed in (\ref{alg:simulation}) and starting the random number generators form the same seed, they will look the same for every $r$ and $t$ as long as their product $rt$ is the same.
This is true for most of the substitution models and implies that typical phylogenetic estimation focuses on inferring the product parameter $\theta=rt$, or genetic distance.
%PL: perhaps mention that the section on molecular clocks (cfr supra) discusses the conditions for which we can start separating time and rate

@@

%PL: why not use a simpler tree for fewer taxa, not necessarily with dated tips, where you indicate the terms you describe. For example the sample tree for figure 1.9, but without node and branch numbering, and taxon 1, taxon 2,.. as tip labels

%---TREE CONCEPT---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{treeconcept} 
\caption{
{ \footnotesize 
{\bf A phylogenetic tree showing evolutionary relationship.} Character sequences are observed at the tips of the tree (black dots); sequence characters at the internal nodes are treated as missing data. 
Branch lengths are estimated in units of genetic distance, or in this case, real time.
} % END: footnotesize
}
\label{fig:treeconcept}
\end{figure}

>>>>

%---SUBSTITUTION MODEL---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{substitution} 
\caption{
{ \footnotesize 
{\bf  Graphical representation of a nucleotide substitution model.} 
Markov substitution model with 6 parameters describing the symmetric transition rates.
} % END: footnotesize
}
\label{fig:substitution}
\end{figure}

@@

\subsection{A simple example: the Jukes and Cantor substitution model\label{sub:jc69}}

@@

\noindent
Second, the rate and time in Equation~(\ref{eq:jc69Finite}) occur as a product, which means that rate and time are confounded.
If we were to simulate two sequences, using the algorithm listed in (\ref{alg:simulation}) and starting the random number generators form the same seed, they will look the same for every $r$ and $t$ as long as their product $rt$ is the same.
This is true for most of the substitution models and implies that typical phylogenetic estimation focuses on inferring the product parameter $\theta=rt$, or genetic distance.
In Subsection~\ref{sub:clocks} we discuss how, under certain conditions, the time and substitution rate can be separated.

@@

%PL: why not use a simpler tree for fewer taxa, not necessarily with dated tips, where you indicate the terms you describe. For example the sample tree for figure 1.10, but without node and branch numbering, and taxon 1, taxon 2,.. as tip labels

%---TREE CONCEPT---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{treeconcept} 
\caption{
{ \footnotesize 
{\bf A phylogenetic tree showing evolutionary relationship.} Character sequences are observed at the tips of the tree (black dots); sequence characters at the internal nodes are treated as missing data (outlined dots). 
Branch lengths can be estimated in units of genetic distance, or as in this case, real time.
} % END: footnotesize
}
\label{fig:treeconcept}
\end{figure}
